
##Libraries and initial directory

library(lubridate)

##General options
```{r setoptions}
opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data

```{r load}
data<-read.csv("activity.csv")
data$date<-as.Date(data$date)
```

## What is mean total number of steps taken per day?

```{r steps_daily}
sum_steps_d<-sapply(split(data,as.factor(data$date)),function(x) sum(x$steps,na.rm=T))
hist(sum_steps_d,
	xlab="number of daily steps",
	ylab="number of days",
	labels=T,
	main="histogram of number of steps per day",
	col="blue")
mean_steps<-mean(sum_steps_d)
median_steps<-median(sum_steps_d)
cat("mean of the total number of steps taken per day:\n");mean_steps
cat("median of the total number of steps taken per day:\n");median_steps
```

## What is the average daily activity pattern?

```{r steps_interval}
ave_steps_i<-sapply(split(data,as.factor(data$interval)),function(x) mean(x$steps,na.rm=T))
plot(x=names(ave_steps_i),y=ave_steps_i,type="l",
	xlab="daily interval",
	ylab="average of number of steps",
	col="purple",
	lwd=8)

max_i<-ave_steps_i[ave_steps_i==max(ave_steps_i)]
cat("5 min interval with the maximum number of steps is:\n");names(max_i)
cat("5 min interval with the maximum number of steps is:\n");cat(names(max_i));cat("\n")
```

## Imputing missing values

```{r nas}
cat("The number of NAs on each variable is the following\n");apply(data,2,function(x) sum(is.na(x)))

#NA imputation
cat("NA imputation\n")
ave_steps_i<-sapply(split(data,as.factor(data$interval)),function(x) mean(x$steps,na.rm=T))
ave_steps_i<-data.frame(interval=as.integer(names(ave_steps_i)),num_steps=ave_steps_i)
data<-merge(data,ave_steps_i,by="interval")
data[is.na(data$steps),]$steps<-data[is.na(data$steps),]$num_steps
data$num_steps<-NULL
cat("The number of NAs on each variable is now\n");apply(data,2,function(x) sum(is.na(x)))

#Recalculation of point 1 without NAs
cat("Recalculations of point 2, histogram of total number of steps per day, mean and median but now without NAs\n")
sum_steps_d<-sapply(split(data,as.factor(data$date)),function(x) sum(x$steps,na.rm=T))
hist(sum_steps_d,
	xlab="number of daily steps",
	ylab="number of days",
	labels=T,
	main="histogram of number of steps per day",
	col="blue")
mean_steps<-mean(sum_steps_d)
median_steps<-median(sum_steps_d)
cat("mean of the total number of steps taken per day:\n");mean_steps
cat("median of the total number of steps taken per day:\n");median_steps
```

**New values differs from the calculated on point 2.**
**Number of steps per days is greater than in point 2**

## Are there differences in activity patterns between weekdays and weekends?

```{r weekends}
#Creation of a variable differentiation weekday and weekday

data$day_type<-"weekday"
data$day_type[weekdays(data$date)%in%weekdays(as.Date(c("2012-12-02","2012-12-01")))]<-"weekend"

#Average number of steps per interval depending on weekday/weekend
data_wd<-subset(data,day_type=="weekday");data_wd$interval<-as.factor(data_wd$interval)
data_we<-subset(data,day_type=="weekend");data_we$interval<-as.factor(data_we$interval)
ave_steps_i_wd<-sapply(split(data_wd,data_wd$interval),function(x) mean(x$steps))
ave_steps_i_we<-sapply(split(data_we,data_we$interval),function(x) mean(x$steps))

#plot
par(mfrow=c(2,1))
plot(x=names(ave_steps_i_wd),y=ave_steps_i_wd,type="l",
	xlab="daily interval",
	ylab="average of number of steps",
	col="purple",
	main="weekday",
	lwd=4)
plot(x=names(ave_steps_i_we),y=ave_steps_i_we,type="l",
	xlab="daily interval",
	ylab="average of number of steps",
	col="purple",
	main="weekend",
	lwd=4)
```



